Artificial Ecosystem Selection

### 2019-06-16 R/fluxes.Rmd
Add new data and update analysis

temp <- list.files(path="../Data", pattern="FluxData*", full.names = TRUE)
conc <- sapply(temp, read.csv, stringsAsFactors = FALSE,
                 simplify = FALSE, USE.NAMES = TRUE)
names(conc) <- substr(names(conc), 18, 25)
conc <- bind_rows(conc, .id = "Date")

# Code transfers
conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))

initial_n1 <- length(unique(conc[conc$transfer == 1, ]$JAR))

initial_n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
```

@@

A set of `r initial_n1` replicate ecosystems were generated
with 50 g of soil (sterile potting mix plus inoculum)
and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
These jars were inoculated at 5% by mass and were initially kept at
atmospheric concentrations of methane. Later, they were capped and spiked to
a headspace concentration of approximately 1000 ppm CH4.  After about 8 weeks
of incubation, we finally observed significant (negative) methane fluxes. Jar
9 had the highest flux, but we observed a crack in the jar and so excluded that
data point. From the remaining `r initial_n1 - 1` jars, we selected the top 4 
(jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.

```{r process_data}
# Jar broke so remove # 9 from gen 1
conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]

# Convert clock time to days and set t0 to 0 days
conc <- 
  conc  %>% 
  mutate(days = period_to_seconds(hms(paste0(conc$TIME, ':00'))) / 60 / 60) %>% 
  group_by(Date, JAR) %>% 
  mutate(days = days - min(days))

# Convert ppm to umol
conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))

# Calculate fluxes from lm slopes
flux <- 
  conc %>% 
  group_by(transfer, JAR) %>%
  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
  filter(term == 'days')  %>% 
  select(transfer, JAR, flux_umold = estimate)

# Identify the jars that were used for selection
flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F

all_par <- flux$flux_umold[flux$transfer == '1']
offspr <- flux$flux_umold[flux$transfer == '2']
sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]

@@

ggplot(flux, aes(x = flux_umold, fill = selected)) +
  geom_histogram(aes(fill = interaction(transfer, selected))) +
  facet_grid(transfer ~ .) +
  scale_fill_manual(values = c('black', 'blue', 'red')) +
  geom_vline(xintercept = mean(all_par)) +
  geom_vline(xintercept = mean(offspr), color = 'blue') +
  geom_vline(xintercept = mean(sel_par), color = 'red') 
ggsave('../Figures/transfer12_hist.pdf', width = 5, height = 5)

ggplot(flux, aes(x = transfer, y = flux_umold)) +
  geom_boxplot(aes(color = transfer), outlier.shape = NA) +
  geom_jitter(aes(color = interaction(transfer, selected))) +
  scale_color_manual(values = c('black', 'black', 'red', 'blue', 'blue')) + 
  geom_hline(yintercept = mean(all_par)) +
  geom_hline(yintercept = mean(offspr), color = 'blue') +
  geom_hline(yintercept = mean(sel_par), color = 'red')

ggsave('../Figures/transfer12_box.pdf', width = 5, height = 5)
# Calculate heritability

@@

nsh <- ns_heritability(all_par, offspr, sel_par) 
```

Based on the mean ecosystem-level trait values from the first and second set
of ecosystems, we calculated narrow-sense heritability as $h^2 = R / S$, which
was `r myround(nsh)`. Granted, this is also not a valid estimate, because of the difference
in inoculum size. We will calculate this again between transfer 2 and 3 to get a more
legitimate estimate.

@@

```{r identify_top_jars} 
flux %>% 
  group_by(transfer) %>% 
  top_n(-4, flux_umold)
```

>>>>

temp <- list.files(path="../Data", pattern="FluxData*", full.names = TRUE)
conc <- sapply(temp, read.csv, stringsAsFactors = FALSE,
                 simplify = FALSE, USE.NAMES = TRUE)
names(conc) <- substr(names(conc), 18, 25)
conc <- bind_rows(conc, .id = "Date")
# Code transfers
#conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))
conc$transfer[conc$Date == 20190527] <- 1
conc$transfer[conc$Date != 20190527] <- 2

# Jar broke so remove # 9 from gen 1
conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]

n1 <- length(unique(conc[conc$transfer == 1, ]$JAR))

n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
```

@@

A set of `r n1` replicate ecosystems were generated
with 50 g of soil (sterile potting mix plus inoculum)
and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
These jars were inoculated at 5% by mass and were initially kept at
atmospheric concentrations of methane. Later, they were capped and spiked to
a headspace concentration of approximately 1000 ppm CH4.  After about 8 weeks
of incubation, we finally observed significant (negative) methane fluxes. Jar
9 had the highest flux, but we observed a crack in the jar and so excluded that
data point. From the 'r n1' jars, we selected the top 4 
(jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.

```{r process_data}

# Convert clock time to days and set t0 to 0 days
conc <- 
  conc  %>% 
  mutate(days = as.numeric(difftime(ymd_hms(paste0(DATE, ' ', TIME, ':00')), +
                       ISOdatetime(2019,1,1,0,0,0), unit="days"))) %>% 
  group_by(Date, JAR) %>% 
  mutate(days = days - min(days))

# Convert ppm to umol
conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))
conc

conc$selected[conc$JAR %in% top_jars_gen1 & conc$transfer == '1'] <- T
conc$selected[!(conc$JAR %in% top_jars_gen1) | conc$transfer != '1'] <- F
```

```{r}

# Calculate fluxes from lm slopes
flux <- 
  conc %>% 
  group_by(Date, JAR) %>%
  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
  filter(term == 'days')#  %>% 
  select(Date, JAR, flux_umold = estimate)

flux <- 
  conc %>% 
  group_by(Date, JAR) %>%
  #do(tidy(glm(ch4_umol ~ days, data = ., family = gaussian(link = "log")))) %>% 
    nest() %>% 
mutate(fit_glm = purrr::map(data, ~ glm(ch4_umol ~ days, data = ., family = gaussian(link = "log"))),
         fit_lm = purrr::map(data, ~ lm(ch4_umol ~ days, data = .)))
flux %>% 
    mutate(aov = map2(fit_glm, fit_lm, stats::anova)) %>% 
    pull(aov)

out %>% 
    mutate(aov = map2(fit_nls, fit_lm, stats::anova)) %>% 
    pull(aov) %>% head(2)

flux$transfer[flux$Date == 20190527] <- 1
flux$transfer[flux$Date != 20190527] <- 2
flux_log$transfer[flux_log$Date == 20190527] <- 1
flux_log$transfer[flux_log$Date != 20190527] <- 2

flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F
flux_log$selected[flux_log$JAR %in% top_jars_gen1 & flux_log$transfer == '1'] <- T
flux_log$selected[!(flux_log$JAR %in% top_jars_gen1) | flux_log$transfer != '1'] <- F
unique(flux$Date)
all_par <- flux$flux_umold[flux$Date == '20190527']
offspr1 <- flux$flux_umold[flux$Date == '20190607']
offspr2 <- flux$flux_umold[flux$Date == '20190614']
sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]

all_par_log <- flux_log$flux_umold[flux_log$Date == '20190527']
offspr1_log <- flux_log$flux_umold[flux_log$Date == '20190607']
offspr2_log <- flux_log$flux_umold[flux_log$Date == '20190614']
sel_par_log <- flux_log$flux_umold[flux_log$transfer == '1' & flux_log$selected == T]
```

```{r plot_flux_curves}
ggplot(conc, aes(x = days, y = ch4_umol, group = JAR)) + 
  geom_point() + 
  facet_wrap(~ Date, scales = "free_x")  + 
    geom_smooth(method = "glm", se = F, 
        method.args = list(family = gaussian(link = "log")))
```

@@

ggplot(flux, aes(x = flux_umold)) +
  geom_histogram(aes(fill = selected)) +
  facet_grid(Date ~ .) +
  scale_fill_manual(values = c('black', 'red')) +
  geom_vline(xintercept = mean(all_par)) +
  geom_vline(xintercept = mean(offspr1), color = 'blue') +
  geom_vline(xintercept = mean(offspr2), color = 'darkblue') +
  geom_vline(xintercept = mean(sel_par), color = 'red')

ggplot(flux, aes(x = Date, y = flux_umold)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color =selected)) +
  scale_color_manual(values = c('black', 'red')) + 
  geom_hline(yintercept = mean(all_par)) +
  geom_hline(yintercept = mean(offspr1), color = 'blue') +
  geom_hline(yintercept = mean(offspr2), color = 'darkblue') +
  geom_hline(yintercept = mean(sel_par), color = 'red')

ggplot(flux_log, aes(x = flux_umold)) +
  geom_histogram(aes(fill = selected)) +
  facet_grid(Date ~ .) +
  scale_fill_manual(values = c('black', 'red')) +
  geom_vline(xintercept = mean(all_par_log)) +
  geom_vline(xintercept = mean(offspr1_log), color = 'blue') +
  geom_vline(xintercept = mean(offspr2_log), color = 'darkblue') +
  geom_vline(xintercept = mean(sel_par_log), color = 'red')

ggplot(flux_log, aes(x = Date, y = flux_umold)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color =selected)) +
  scale_color_manual(values = c('black', 'red')) + 
  geom_hline(yintercept = mean(all_par_log)) +
  geom_hline(yintercept = mean(offspr1_log), color = 'blue') +
  geom_hline(yintercept = mean(offspr2_log), color = 'darkblue') +
  geom_hline(yintercept = mean(sel_par_log), color = 'red') 
```

@@

nsh1 <- ns_heritability(all_par, offspr1, sel_par) 
nsh2 <- ns_heritability(all_par, offspr2, sel_par) 
nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par_log)
nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par_log)
```

Based on the mean ecosystem-level trait values from the first and second set
of ecosystems, we calculated narrow-sense heritability as $h^2 = R / S$, which
was `r myround(nsh)` for the linear model and `r myround(nsh_log)` for the exponential
model. Granted, this is also not a valid estimate, because of the difference
in inoculum size. We will calculate this again between transfer 2 and 3 to get a more
legitimate estimate.

@@

```{r identify_top_jars} 
flux %>% 
  group_by(Date) %>% 
  top_n(-4, flux_umold)
flux_log %>% 
  group_by(Date) %>% 
  top_n(-4, flux_umold)
```
